---
title: "Data Scraping with rvest"
author: "Math 271"
date: "Spring 22"
output: 
    html_document:
        css: lab.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(rvest)
library(lubridate)
```

## Table Scraping

Web pages often contain tables with data:

https://www.timeanddate.com/weather/@5850911/historic

Using a computer to automatically extract data from human-readable web pages is known as _scraping_. There are some possible ethical and terms-of-use issues involved in scraping, especially if you try to do it on a large scale on many pages, but if you're just looking at one or two tables from a public website, there is probably nothing to worry about. 
The `rvest` library has functions for dealing to webpages. We begin with the function `read_html`, which fetches and parses a webpage.

```{r}
raw_html <- read_html("https://www.timeanddate.com/weather/@5850911/historic")
```

google: "mauna kea weather records"

## CSS selectors

A html document contains a nested tree of _elements_ that make up what is known as a _DOM_ (Document Object Model). An element looks like

    <element class="class1 class2" id="uniqueID">...element-content...</element>
    
Visit the `timeanddate.com` link above in your browser, then right click on the data table and open the _Developer Tools_ by choosing the _Inspect_ menu item. This website explains more about the [Dev Tools](https://developer.mozilla.org/en-US/docs/Learn/Common_questions/What_are_browser_developer_tools) 

Make sure the devtools window has the "Elements" tab selected, where the DOM is displayed. You can expand and contract the DOM tree using the triangles. Selecting and hovering over elements in the Dev Tools will highlight the element in the page view.

Our goal is to be able to identify the elements containing the data we want to extract. Data is often stored in a `table` element. Locate the Elements with the LOTR data tables.

We will extract the data by constructing a [CSS selector](https://www.w3schools.com/cssref/css_selectors.asp) that will grab only the data from the webpage.

Here's a game about writing CSS selectors: http://flukeout.github.io/

In this case we are lucky and the data table has a unique id, `id="wt-his"` which we can use to easily extract the table. The CSS selector for a known id is `#id`.

We use the `rvest` functions `html_node` to extract the table from the page, and `html_table` to process the table into a data frame.

```{r}
raw_html %>% html_node("#wt-his") %>% html_table(header=FALSE)
```
Now the challenge of wrangling this data into shape begins.


```{r, include=FALSE}
raw_html %>% 
    html_node("#wt-his") %>% 
    html_table(header = FALSE) %>% 
    set_colnames(.,.[2,]) %>% 
    repair_names() %>% 
    slice(3:(n()-1)) %>%
    separate(Time,c("Time", "Date"), "(?<=[ap]m)") %>% 
    mutate(Date=na_if(Date, "")) %>% 
    fill(Date) %>% 
    select(-V1,-V2) %>% 
    # mutate(Temp=str_extract(Temp,"\\d+"),
    #        Humidity=str_extract(Humidity,"\\d+"), 
    #        Barome) %>% 
    mutate(across(Temp|Wind:Visibility, str_extract, "\\d+")) %>% 
    mutate(dt=parse_date_time(paste(Time, Date),"hMpamd", "Pacific/Honolulu")) %>% 
    type_convert()
```


## Stylin' with CSS

The CSS selectors we used to extract the data tables is the same language we use to apply style to elements in our Markdown documents. If you inspect the output document you will find that the section headers are `<h2>` elements. Let's change their color to `#b60101`, a nice dark UHH-red color, and put a line under them.

```{css}
h2 {
    color:  #b60101;
    border-bottom: 1px solid;
}
```

Most medium-to-large organizations have a _Style Guide_ with official recommendations for things like color schemes. UH Hilo's style guide is found at https://hilo.hawaii.edu/assets/styleguide/. If you want to use your own colors, https://coolors.co/ is a decent place to start building a color palette that won't hurt your readers' eyes.

A full guide to what you can achieve with CSS styling can be found here: https://www.w3schools.com/css/

Notice that the _code chunk_ above is not a `{r}` chunk, but a `{css}` chunk. This tells the Knitr that this is CSS code, not R code, and it will handle it differently. Setting the chunk option `echo=FALSE` is a common thing to do with CSS chunks.

## Other Language Chunks

This section is mostly for informational purposes, I've never used these features in practice.

The Knitr markdown processor recognizes a few other languagues, below is an example of a `{python}` language chunk. Python chunks are processed similarly to R chunks, with the commands executed and results inserted in the output document.

```{python, python.reticulate=FALSE}
import sys
print "Hello from Python version", sys.version
```

For html output formats, the `{js}` language is available, this code is different than R or Python chunks, in that it is inserted directly into the webpage, and is run on the browser viewing the page rather than on the computer that is Knitting the document.

```{js}
$("<p>This content added in browser by JavaScript.</p>").appendTo(".main-container")
```

